---
title: 'Problem Set 2: Omitted Variable Bias and Fixed Effects'
author: "Claire Duquennois"
output:
  pdf_document: default
  html_document:
    number_sections: yes
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

***NAME: Benjamin Houck***

**Empirical Analysis using Data from Washington (2008, AER)**

This exercise, like PS1, also uses data from Ebonya Washington's paper, "Female Socialization: How Daughters Affect their Legislator Father's voting on Women's Issues," published in the *American Economic Review* in 2008. This paper studies whether having a daughter affects legislator's voting on women's issues.

**Submission instructions:**

1)  Knit your assignment in PDF.

2)  **Assignment shell extra credit:** You will receive a little extra credit if your answers line up correctly with the answer positions of the template on gradescope. For this to work

-   **Work by putting your answers directly into this R markdown file that is pre-formatted for you!**

-   Make sure you have ONE question and answer per page (this is how the template is structured and allows gradescope to easily find your answers), unless the question indicated that the answer will require two pages because of large tables.

-   Your final PDF should be 18 pages.

-   You can insert needed page breaks as illustrated below

-   Make sure you do not "print" the data. If you change the data, make sure you store with a structure like: newname\<-modification(olddata). If you only type modification(olddata), R will display the data rather than saving your modifications

3)  Upload your assignment PDF to gradescope.

\pagebreak

# Set up and opening the data

## Question 1.1:

**Load the `basic.dta` file like you did for PS1 and call all the packages you will be using with the `library` function. The packages you will need are `haven`, `lfe`, `dplyr`, and `stargazer`.**

**Code:**

```{r data_setup}
library(dplyr)
library(lfe)
library(stargazer)
library(haven)

basic <- read_dta("basic.dta")

```

\pagebreak

# Cleaning the data

## Question 2.1:

**Like in PS1, restrict your data to observations from the 105th congress and keep only the variables listed in the table below. Here too, make sure your final dataset is a data frame.**

| Name    | Description                                                              |
|--------------------|----------------------------------------------------|
| aauw    | AAUW score                                                               |
| totchi  | Total number of children                                                 |
| ngirls  | Number of daughters                                                      |
| party   | Political party. Democrats if 1, Republicans if 2, and Independent if 3. |
| female  | Female dummy variable                                                    |
| white   | White dummy variable                                                     |
| srvlng  | Years of service                                                         |
| age     | Age                                                                      |
| demvote | State democratic vote share in most recent presidential election         |
| rgroup  | religious group                                                          |
| region  | region                                                                   |
| name    | representative's name                                                    |

**You can find the detailed description of each variable in the original paper. The main variable in this analysis is `AAUW`, a score created by the American Association of University Women (AAUW). For each congress, AAUW selects pieces of legislation in the areas of education, equality, and reproductive rights. The AAUW keeps track of how each legislator voted on these pieces of legislation and whether their vote aligned with the AAUW’s position. The legislator’s score is equal to the proportion of these votes made in agreement with the AAUW.**

Hint: You can use the `filter` and `select` functions we discussed in Slides2_IntroR.

**Code**

```{r filtering, include=TRUE}
basic_105 <-basic %>%
  filter(congress==105)
filtered_105 <-basic_105 %>% 
  select(aauw,totchi,ngirls,party,female,white,srvlng,age,demvote,rgroup,region,name)

is(filtered_105)

data_df = data.frame(as.matrix(filtered_105))

is(data_df)

```

\pagebreak

# Analysis

## Question 3.1:

**Estimate the following linear regression models using the `felm` command (part of the lfe package). Report your regression results in a formatted table using `stargazer`. Report robust standard errors in your table.**

$$
\begin{aligned}
 \text{Model 1: }aauw_i&=\beta_0+\beta_1ngirls_i+\epsilon_i\\
 \text{Model 2: }aauw_i&=\beta_0+\beta_1ngirls_i+\beta_2totchi+\epsilon_i\\
\end{aligned}
$$ Hints: If you want RMarkdown to display your outputted table, include the code `results = "asis"` in the chunk header. This is true for all chunks that output a formatted table. In the stargazer command, you will want to specify the format of the table by including the code `type="latex"` for a pdf output. If you have trouble knitting to PDF, try installing MikTex (<https://miktex.org/download>). You will need to specify `se = list(reg1$rse, reg2$rse)` in stargazer, like the example in Slides3_RegressionReview.

**Code:**

```{r preliminary regressions, results="asis"}

data_df$ngirls <- as.numeric(data_df$ngirls)  # 
data_df$totchi <- as.numeric(data_df$totchi)

model1 = felm(aauw ~ ngirls, data = data_df) # modeling ngirls
model2 = felm(aauw ~ ngirls + totchi, data = data_df)


# Report the results using stargazer with robust standard errors
stargazer(model1, model2, 
          se = list(model1$rse, model2$rse),
          type = "latex",
          title = "Regression Results",
          dep.var.labels = "AAUW",
          covariate.labels = c("Number of Girls", "Total Children"),
          omit.stat = c("f", "ser"),
          keep = c("ngirls", "totchi"),
          header = FALSE)


```

\pagebreak

## Question 3.2:

**Compare the estimates of** $\beta_1$ across the two specifications. Why does our estimate of $\beta_1$ changes so much? Which control variable is particularly important and why?

**Answer:**

$\beta_1$ changes so much because we we do not control for total children. We see in $\beta_2$ from the second regression there is a large negative correlation between number of children and AAUW score. This is likely because large families is a core conservative value and therefore would likely have negative correlation with AAUW scores. This demonstrates an example of downward bias with a negative OVB from number of children.

\pagebreak

## Question 3.3:

**Consider the second specification which controls for** $totchi_i$. Conditional on the number of children, do you think $ngirls_i$ is plausibly exogenous? What is the identifying assumption, i.e. the assumption that must hold for $\beta_1$ to be interpreted as a causal estimate? What evidence does Washington give to support this assumption?

**Answer:**

Yes, I believe \(ngirls_i\) is plausibly exogenous when controlling for \(totchi_i\). The identifying assumption is that the number of daughters is as good as random, meaning that any variation in \(ngirls_i\) does not correlate with unobserved factors affecting the outcome. For \(\beta_1\) to be interpreted as a causal estimate, we must assume that the sex composition of children does not systematically relate to parental characteristics or socio-economic status. Washington supports this assumption by providing evidence showing that the birth of girls versus boys appears random and independent of these confounding factors, thereby reinforcing the argument for exogeneity.

\pagebreak

# Fixed Effects:

## Question 4.1 (2 pages):

**Equation 1 from Washington's paper is a little bit different from the equations you have estimated so far. Estimate the three models specified below (where** $\gamma_i$ is a fixed effect for the number of children). Present your results in a table. Use robust standard errors.

$$
\begin{aligned}
  \text{Model 1: }aauw_i&=\beta_0+\beta_1ngirls_i+\beta_2totchi+\epsilon_i\\
   \text{Model 2: }aauw_i&=\beta_0+\beta_1ngirls_i+\beta_2chi1+...+\beta_{10}chi10 +\epsilon_i\\
    \text{Model 3: } aauw_i&=\beta_0+\beta_1ngirls_i+\gamma_i+\epsilon_i\\
\end{aligned}
$$

Hints:

-   you will need to generate the dummy variables for the second equation or code it as `factor(totchi)`.

-   For the third equation, the `felm` function allows you to specify fixed effects as we saw in class.

**Code:**
```{r finalized regressions, results="asis"}
model4  <-felm(aauw ~ ngirls + totchi, data=data_df)
model5 <-felm(aauw ~ ngirls + factor(totchi), data=data_df)
model6 <- felm(aauw ~ ngirls | totchi, data = data_df)

stargazer(model4, model5, model6, 
          type = "latex",
          title = "Regression Results",
          dep.var.labels = "AAUW",
          covariate.labels = c("Number of Girls"),
          omit.stat = c("f", "ser"),
          header = FALSE)
```
\pagebreak

## Question 4.2:

**Explain the difference between the three models.**

**Answer:**

They show a similar value for $%\beta_1%$ in all three models. Total children is controlled for in all models through different representations. The first model shows the lowest $R^2$ value but provides a straight forward regression. The second regression shows a higher $R^2$ value however provides all data points of total kids as dummy variables, making the regression larger than it needs to be, due to the size of the data set. The third regression shows the same high $R^2$ but is shortened utilizing the fixed effect rather than the dummy variables. The third regression provides the easiest interpretation for the reader with the highest $R^2$ value.


\pagebreak

## Question 4.3 (2 pages):

**Reproduce the EXACT results presented in column 2 of table 2 from Washington's paper. To do this you will need to first build three variables:** $age^2$ and $srvlng^2$ and $repub_i$, an indicator set to 1 if the representative is republican and 0 otherwise. Then estimate the following specification, where $\gamma_i$ is a fixed effect for total children, $\phi_i$ is a fixed effect for religious group, and $\lambda_i$ is a fixed effect for region:

$$
\text{Model A: }aauw_i=\beta_0+\beta_1ngirls_i+female_i+white_i+repub_i+age_i+age^2_i+srvlng_i+srvlng^2_i+demvote_i+\gamma_i+\phi_i+\lambda_i+\epsilon_i
$$ Hint:

-   Because you are using a fixed effect for religion, stargazer will not print the religion coefficients. That is ok, those coefficients are still being estimated in the background of the R program.

-   Do not use robust standard errors as she does not in this table.

**Code:**
```{r table1, results="asis"}
data_df$age <- as.numeric(data_df$age)
data_df$srvlng <- as.numeric(data_df$srvlng)
data_df$demvote <- as.numeric(data_df$demvote)


# Create the squared terms separately (before model estimation)
data_df$age_sq <- data_df$age^2
data_df$srvlng_sq <- data_df$srvlng^2
#create dummy variable for republican identity
repub = ifelse(data_df$party == 2, 1, 0)

table1 = felm(aauw ~ngirls+female+white+repub+age+age_sq+srvlng+srvlng_sq+demvote|totchi+rgroup+region,(data=data_df))

stargazer(table1, 
          type = "latex",
          title = "Regression Results",
          dep.var.labels = "AAUW",
          covariate.labels = c("Number of Girls","Female","White","Republican","Age","Age Squared","Service Length","Service Length Squared","Democratic vote share in district \\\\ \\hspace{0.5cm}(most recent presidential election)"),
          omit.stat = c("f","ser"),
          header = FALSE,
          digits = 2)
```

\pagebreak

## Question 4.4:

**Explain what the region fixed effects are controlling for.**

**Answer:**
The region fixed effects control for the regional effects on AAUW voting. It is generally known southern states lean more to the right and northern states to lean more to the left, specifically on abortion and women's rights. Recognizing this, politicians may be swayed to appeal to their voting base. Controlling for this allows the region and regional bias to not affect the coefficients. 

\pagebreak

## Question 4.5:

**Reload the data and this time keep observations from all of the four congresses. Add the three variables you built for question 4.3 to this data set**

**Code:**

```{r table2, results="asis"}

basic <- read_dta("basic.dta")

filtered_full <-basic %>% 
  select(aauw,totchi,ngirls,party,female,white,srvlng,age,demvote,rgroup,region,,congress,name)
#reformatting as data frame
data_df_full = data.frame(as.matrix(filtered_full))

#ensure variables are numeric rather than dummy
data_df_full$ngirls <- as.numeric(data_df_full$ngirls)
data_df_full$age <- as.numeric(data_df_full$age)
data_df_full$srvlng <- as.numeric(data_df_full$srvlng)
data_df_full$demvote <- as.numeric(data_df_full$demvote)

# Create the squared terms separately (before model estimation)
data_df_full$age_sq <- data_df_full$age^2
data_df_full$srvlng_sq <- data_df_full$srvlng^2
#create dummy variable for republican identity
repub = ifelse(data_df_full$party == 2, 1, 0)
```


\pagebreak

## Question 4.6 (2 pages):

**Because we have data for four congress sessions, we may be able to see how an individual congress person's voting patterns change as the number of daughters they have changes. Estimate model A with the addition of `congress` and `name` fixed effects. Present your results in a table.**

**Code:**
```{r table3, results="asis"}

data_df_full$congress <- as.numeric(data_df_full$congress)

table_3 = felm(aauw ~ngirls+female+white+repub+age+age_sq+srvlng+srvlng_sq+demvote|totchi+rgroup+region+congress+name,(data=data_df_full))

stargazer(table_3, 
          type = "latex",
          title = "Regression Results",
          dep.var.labels = "AAUW",
          covariate.labels = c("Number of Girls", 
                               "Female", 
                               "White", 
                               "Republican", 
                               "Age", 
                               "Age Squared", 
                               "Service Length", 
                               "Service Length Squared", 
                               "Democratic vote share in district \\\\ \\hspace{0.5cm}(most recent presidential election)"),
          omit.stat = c("f", "ser"),
          header = FALSE,
          digits = 2)


```
\pagebreak

## Question 4.7:

**How does this estimate compare to your estimate in question 4.3? Why are the standard errors so much bigger? Why doesn't Washington use this approach in her paper?**

**Answer:**

This estimate shows high standard errors because there is multicollinearity between variables. Controlling for congress and name will omit many data points by factoring them into fixed effects, such that every other variable will only account for changes in variables. For example a politician who does not have a daughter during their terms will be omitted from factoring into coefficient $%\beta_1%$ eliminating data and creating a higher standard error. The same is true for every other independent variable as they will be factored into congress or name unless there is a change over their terms.

\pagebreak

## Question 4.8:

**Why are you not able to generate a coefficient for** $female_i$ or $white_i$?

**Answer:**

There are no politicians who changed sex or race during their terms therefore these variables are entirely encompassed by fixed effects.

\pagebreak

## Question 4.9:

**You are able to generate an estimate for** $repub_i$. What does this imply?

**Answer:**

This implies there was at least one example of a politician who changed parties between 1997 and 2004. We control for name and congress, therefore those who maintained their party status will have this attribute accounted for in fixed effects. The variable Republican therefore shows how changing party identification will affect a politician's voting scores. This explains why it is a negative affect, as Republican politicians tend to not vote for women's rights compared to Democrats. Therefore those who changed parties from Independent or Democrat to Republican will have a lower AAUW score compared to when they identified as an Independent or Republican.